---
title: "Data Analysis for pathogenicity across variation in microbial virulence and host resistance"
author: "Anjali Gupta"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all required libraries

If one or more packages is not already installed for you, you can install it using ```install.packages("PackageName")```


```{r message=FALSE, warning=FALSE}

library(readr)
library(dplyr)
library(stringr)
library(survival)
library(ggplot2)
library(survminer)
library(wesanderson)
library(ggpubr)
library(coxme)
library(lme4)
library(corrplot)
library(RColorBrewer)
library(ggrepel)
library(reshape2)

```

## Set your working directory and import dataset

Set your working directory to the folder where your dataset is present

The dataset can be downloaded from [Main Data](https://github.com/anjaligupta1210/Pathogen-Potential/blob/main/Data.csv) and [Bacterial Load Data](https://github.com/anjaligupta1210/Pathogen-Potential/blob/main/PathogenPotential_BacterialLoad_AG.csv)

```{r message=FALSE, warning=FALSE}
setwd("/Users/anjaligupta/Documents/GitHub/Pathogen-Potential")

Data <- read_csv("Data.csv")

LoadData <- read_csv("PathogenPotential_BacterialLoad_AG.csv")

```

## Data processing and transformation to survival format

Entering Dpt host genotype information

```{r message=FALSE, warning=FALSE}
Data <- Data %>%
  mutate(Genotype = recode(Genotype,
                           "1G" = "DptNull",
                           "1N" = "DptR",
                           "8N" = "DptS"))
```

Converting survival data to 0/1 format

```{r message=FALSE, warning=FALSE}
#Create an empty dataframe Data_1 to store the survival information

Data_1 <- data.frame(Vial_ID = character(),
                     Genotype = character(),
                     Treatment = character(),
                     OD_group = character(),
                     Date = character(),
                     Day_Dead = numeric(),
                     censor = numeric())


#Convert data to 0/1 format

for(i in 1:length(Data$Vial_ID)){
  temp=data.frame(Data[i,])
  
  for(j in seq(10,22,2)){
    k = temp[1,j-2]-temp[1,j]
    if(k>0){
      for(kk in 1:k){
        Data_1[length(Data_1$Vial_ID)+1,]=c(temp[1,1],
                                            as.character(temp[1,2]),
                                            as.character(temp[1,3]),
                                            temp[1,5],
                                            as.character(temp[1,6]),
                                            j/2 - 4,
                                            1)
      }    
    } 
  }
  
  for(j in 8){
    k = temp[1,j]
    if(k>0){
      for(kk in 1:k){
        Data_1[length(Data_1$Vial_ID)+1,]=c(temp[1,1],
                                            as.character(temp[1,2]),
                                            as.character(temp[1,3]),
                                            temp[1,5],
                                            as.character(temp[1,6]),
                                            j/2 - 4,
                                            0)
      }    
    } 
  }
  
  for(j in 22){
    k = temp[1,j]
    if(k>0){
      for(kk in 1:k){
        Data_1[length(Data_1$Vial_ID)+1,]=c(temp[1,1],
                                            as.character(temp[1,2]),
                                            as.character(temp[1,3]),
                                            temp[1,5],
                                            as.character(temp[1,6]),
                                            j/2 - 4,
                                            0)
      }
    }
  }
}


#Converting Data_1 columns from character to numeric/factors

Data_1$Day_Dead = as.numeric(Data_1$Day_Dead)
Data_1$censor = as.numeric(Data_1$censor)

Data_1$Genotype <- as.factor(Data_1$Genotype)
Data_1$Treatment <- as.factor(Data_1$Treatment)
Data_1$Date <- as.factor(Data_1$Date)
Data_1$OD_group <- as.factor(Data_1$OD_group)

#Data_1 is the dataframe containing only survival data 
```

Calculating fraction symptomatic

Dead individuals on any day are assumed to be symptomatic

```{r message=FALSE, warning=FALSE}
Data$Fs_Day1 <- (Data$Day0_Alive - Data$Day1_Climbing) / Data$Day0_Alive
Data$Fs_Day2 <- (Data$Day0_Alive - Data$Day2_Climbing) / Data$Day0_Alive
Data$Fs_Day3 <- (Data$Day0_Alive - Data$Day3_Climbing) / Data$Day0_Alive
Data$Fs_Day4 <- (Data$Day0_Alive - Data$Day4_Climbing) / Data$Day0_Alive
Data$Fs_Day5 <- (Data$Day0_Alive - Data$Day5_Climbing) / Data$Day0_Alive
Data$Fs_Day6 <- (Data$Day0_Alive - Data$Day6_Climbing) / Data$Day0_Alive
Data$Fs_Day7 <- (Data$Day0_Alive - Data$Day7_Climbing) / Data$Day0_Alive


#Creating a dataframe Fs for fraction symptomatic by transforming data
Fs <- melt(Data[,c(1:3,5,6,24:30)], id=c("Vial_ID",
                                         "Genotype",
                                         "Treatment",
                                         "OD_group",
                                         "Date"))
colnames(Fs)[7] <- "Fs"


Fs[c('F', 'Day')] <- str_split_fixed(Fs$variable, '_Day', 2)
Fs <- Fs[,c(1:5,7,9)]
```

Calculating Pathogen Potential

PP = (Fs/I) * (10^M)

```{r message=FALSE, warning=FALSE}

Data$PP_Day1 <- (Data$Fs_Day1/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day1_Alive)/Data$Day0_Alive))

Data$PP_Day2 <- (Data$Fs_Day2/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day2_Alive)/Data$Day0_Alive))

Data$PP_Day3 <- (Data$Fs_Day3/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day3_Alive)/Data$Day0_Alive))

Data$PP_Day4 <- (Data$Fs_Day4/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day4_Alive)/Data$Day0_Alive))

Data$PP_Day5 <- (Data$Fs_Day5/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day5_Alive)/Data$Day0_Alive))

Data$PP_Day6 <- (Data$Fs_Day6/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day6_Alive)/Data$Day0_Alive))

Data$PP_Day7 <- (Data$Fs_Day7/exp(Data$OD_group)) * 
  (10 ^ ((Data$Day0_Alive - Data$Day7_Alive)/Data$Day0_Alive))

#Creating a dataframe PP for pathogen potential by transforming data
PP <- melt(Data[,c(1:3,5,6,31:37)], id=c("Vial_ID",
                                         "Genotype",
                                         "Treatment",
                                         "OD_group",
                                         "Date"))
colnames(PP)[7] <- "PP"

PP[c('p', 'Day')] <- str_split_fixed(PP$variable, '_Day', 2)

PP <- PP[,c(1:5,7,9)]
```

Calculating proportion survival 

```{r message=FALSE, warning=FALSE}
#Proportion survival on Day X
Data$Prop_Day0_Alive <- Data$Day0_Alive/Data$Day0_Alive
Data$Prop_Day1_Alive <- Data$Day1_Alive/Data$Day0_Alive
Data$Prop_Day2_Alive <- Data$Day2_Alive/Data$Day0_Alive
Data$Prop_Day3_Alive <- Data$Day3_Alive/Data$Day0_Alive
Data$Prop_Day4_Alive <- Data$Day4_Alive/Data$Day0_Alive
Data$Prop_Day5_Alive <- Data$Day5_Alive/Data$Day0_Alive
Data$Prop_Day6_Alive <- Data$Day6_Alive/Data$Day0_Alive
Data$Prop_Day7_Alive <- Data$Day7_Alive/Data$Day0_Alive
```

Calculating cumulative risk scores <https://stat.ethz.ch/R-manual/R-devel/library/survival/html/predict.coxph.html>

```{r message=FALSE, warning=FALSE}
Data_Risk <- subset(Data_1, !Day_Dead==0)
cox_model <- coxph(Surv(Day_Dead, censor) ~ as.factor(Vial_ID) +
                     Genotype + 
                     Treatment + 
                     OD_group, 
                   data = Data_Risk)

# Compute the risk scores
risk_scores <- predict(cox_model, type = 'risk')

#Add it to a dataframe
Data_Risk$Risk_Score <- risk_scores

#group by vial id and day to assign cumulative scores

Data_1_2 <- Data_Risk %>%
  group_by(Vial_ID,Genotype,Treatment,OD_group) %>%
  summarise(Vial_ID=Vial_ID,
            Genotype=Genotype,
            Treatment=Treatment,
            OD_group=OD_group,
            Risk_Score=sum(Risk_Score)) %>%
  distinct()

#Add risk scores to the main dataframe
Data <- merge(Data, Data_1_2)
```

Calculating PP_T & Fs_T & Fs/T 

T=median survival day

```{r message=FALSE, warning=FALSE}
#PP_T Fs_T Fs/T

#Estimating mean survival day
Surv <- melt(Data[,c(1:4,6,39:45)], id=c("Vial_ID",
                                         "Genotype",
                                         "Treatment",
                                         "OD_group",
                                         "Date"))
colnames(Surv)[7] <- "Prop_alive"

Surv[c('p', 'Day')] <- str_split_fixed(Surv$variable, '_Alive', 2)
Surv[c('p', 'Day')] <- str_split_fixed(Surv$p, 'Prop_Day', 2)

Surv <- Surv[,c(1:5,7,9)]

Surv$Day_T <- 7

for(i in 1:length(Surv$Vial_ID)){
  temp=data.frame(Surv[i,])
  
  for(j in 6) {
    k=temp[1,j]
    
    if(k<=0.5) {
      Surv$Day_T[i] = Surv$Day[i]
    }
  }
}


Surv_1 <- Surv %>%
  group_by(Vial_ID,Genotype,Treatment,OD_group) %>%
  summarise(Vial_ID=Vial_ID,
            Genotype=Genotype,
            Treatment=Treatment,
            OD_group=OD_group,
            Day_T=min(Day_T)) %>%
  distinct()

Data <- merge(Data, Surv_1)

Data$Fs_T <- NA
Data$PP_T <- NA

#Estimating pathogen potential and fraction symptomatic on mean survival day T

for(i in 1:length(Data$Vial_ID)) {
  temp=data.frame(Data[i,])
  
  j=temp[1,47]
  
  if(j==1) {
    Data$Fs_T[i] = Data$Fs_Day1[i]
    Data$PP_T[i] = Data$PP_Day1[i]
  }
  
  if(j==2) {
    Data$Fs_T[i] = Data$Fs_Day2[i]
    Data$PP_T[i] = Data$PP_Day2[i]
  }
  
  if(j==3) {
    Data$Fs_T[i] = Data$Fs_Day3[i]
    Data$PP_T[i] = Data$PP_Day3[i]
  }
  
  if(j==4) {
    Data$Fs_T[i] = Data$Fs_Day4[i]
    Data$PP_T[i] = Data$PP_Day4[i]
  }
  
  if(j==5) {
    Data$Fs_T[i] = Data$Fs_Day5[i]
    Data$PP_T[i] = Data$PP_Day5[i]
  }
  
  if(j==6) {
    Data$Fs_T[i] = Data$Fs_Day6[i]
    Data$PP_T[i] = Data$PP_Day6[i]
  }
  
  if(j==7) {
    Data$Fs_T[i] = Data$Fs_Day7[i]
    Data$PP_T[i] = Data$PP_Day7[i]
  }
  
}

Data$`Fs_T/T` <- Data$Fs_T/as.numeric(Data$Day_T)
```

## Main Figures

### Figure 1B

```{r message=FALSE, warning=FALSE}
Fig1b <- ggplot(Data, aes(x = as.factor(OD_group),
                 y = Prop_Day4_Alive,
                 pch = as.factor(OD_group),
                 col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.3) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  facet_grid(~Treatment) +
  coord_cartesian(ylim=c(0,1)) +
  theme_classic() +
  labs(x = "OD",
       y = "Proportion alive on day 4",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30))
```

**Figure 1B** - Interval plot of the proportion of flies alive on day 4 post infection for various host genotypes (homozygous serine dpt (wildtype, dptS69), homozygous arginine dpt (dptS69R), and dpt null (Δdpt)), and variable pathogen strains (*P. burhodograneria* Strain D, *P. rettgeri*, *P. sneebia*, *P. alcalifaciens*) with different loads of infective inoculum (OD).

```{r fig.width=30, fig.height=12, message=FALSE, echo=FALSE, warning=FALSE}
Fig1b
```

### Figure 2

```{r message=FALSE, warning=FALSE}

Data_1$Treatment <- factor(Data_1$Treatment, levels = c("P. burhodogranareia strD",
                                                        "P. rettgeri",
                                                        "P. sneebia",
                                                        "P. alcalifaciens"))

fit <- survfit(Surv(Day_Dead, censor) ~ OD_group + Treatment + Genotype,
               data = Data_1)

# Figure 2A
Survival <- ggplot(data = surv_summary(fit), 
                   aes(x = time, 
                       y = surv, 
                       group = interaction(Genotype,
                                           OD_group))) +
  facet_grid(~Treatment) +
  theme_classic() +
  labs(x = "Days post infection",
       y = "Survival") +
  geom_step(aes(color = Genotype, 
                linetype = OD_group),
            size = 1.5) +
  coord_cartesian(xlim = c(0, 7), 
                  ylim = c(0, 1)) +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(linetype = "OD") +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30))

Fs$Treatment <- factor(Fs$Treatment, levels = c("P. burhodogranareia strD",
                                                "P. rettgeri",
                                                "P. sneebia",
                                                "P. alcalifaciens"))
# Figure 2B
FracSick <- ggplot(subset(Fs,
                         Day==4), 
                aes(x = as.factor(OD_group), 
                    y = Fs, 
                    pch = as.factor(OD_group),
                    col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.3) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  facet_grid(~Treatment) +
  theme_classic() +
  coord_cartesian(ylim = c(0,1)) +
  labs(x = "OD",
       y = "Fs on day 4",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30)) +
  guides(color=FALSE)

Data$Treatment <- factor(Data$Treatment, levels = c("P. burhodogranareia strD",
                                                    "P. rettgeri",
                                                    "P. sneebia",
                                                    "P. alcalifaciens"))
# Figure 2C
Risk <- ggplot(Data, 
                aes(x = as.factor(OD_group), 
                    y = Risk_Score,
                    pch = as.factor(OD_group),
                    col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.3) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  facet_grid(~Treatment) +
  theme_classic() +
  labs(x = "OD",
       y = "Risk Score",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30)) +
  guides(color=FALSE)

Tile2 <- ggarrange(FracSick, Risk, nrow = 1, ncol = 2, labels = c("B", "C"), common.legend = TRUE)

Figure2 <- ggarrange(Survival, Tile2, nrow = 2, ncol = 1, common.legend = TRUE, labels = c("A",""))
```

**Figure 2** - Host-based measures of pathogenicity:

* A. Survival curves, 

* B. Fraction symptomatic on Day4, and 

* C. Risk scores for infected flies of various host genotypes (homozygous serine dpt (wildtype, dptS69), homozygous arginine dpt (dptS69R), and dpt null (Δdpt)), and variable pathogen strains (*P. burhodograneria* Strain D, *P. rettgeri*, *P. sneebia*, *P. alcalifaciens*) with different loads of infective inoculum (OD).

```{r fig.width=40, fig.height=18, message=FALSE, echo=FALSE, warning=FALSE}
Figure2
```

### Figure 3

```{r message=FALSE, warning=FALSE}
# Figure 3A
PP_D4 <- ggplot(Data, 
                aes(x = as.factor(OD_group), 
                    y = PP_Day4,
                    pch = as.factor(OD_group),
                    col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.6) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  stat_smooth(se = FALSE, 
              formula = y ~ x, 
              method = "lm") +
  facet_grid(~Treatment) +
  theme_classic() +
  labs(x = "OD",
       y = "PP on day 4",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30))

# Figure 3B
PP_T <- ggplot(Data, 
                aes(x = as.factor(OD_group), 
                    y = PP_T,
                    pch = as.factor(OD_group),
                    col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.6) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  stat_smooth(se = FALSE, 
              formula = y ~ x, 
              method = "lm") +
  facet_grid(~Treatment) +
  theme_classic() +
  labs(x = "OD",
       y = "PP_T",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30))

# Figure 3C
Fs_T <- ggplot(Data, 
                aes(x = as.factor(OD_group), 
                    y = Fs_T,
                    pch = as.factor(OD_group),
                    col = Genotype)) +
  stat_summary(geom = "errorbar", 
               position = position_dodge(0.7), 
               width = 0.6) +
  stat_summary(geom = "point", 
               position = position_dodge(0.7), 
               size = 6) +
  stat_smooth(se = FALSE, 
              formula = y ~ x, 
              method = "lm") +
  facet_grid(~Treatment) +
  theme_classic() +
  labs(x = "OD",
       y = "Fs_T",
       pch="OD") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  theme(legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.title = element_text(size = 30),
        strip.background = element_blank(),
        strip.text = element_text(size = 30, face = "italic"),
        axis.text = element_text(size = 30))

Figure3 <- ggarrange(PP_D4, PP_T, Fs_T, nrow = 3, ncol = 1, common.legend = TRUE, labels = "AUTO")
```

**Figure 3** 

* A. Pathogen potential on Day 4 post-infection, 
* B. Pathogen potential on median survival day (T), and 
* C. Fraction symptomatic on median survival day (T) for infected flies of various host genotypes (homozygous serine dpt (wildtype, dptS69), homozygous arginine dpt (dptS69R), and dpt null (Δdpt)), and variable pathogen strains (*P. burhodograneria* Strain D, *P. rettgeri*, *P. sneebia*, *P. alcalifaciens*) with different loads of infective inoculum (OD).

```{r fig.width=30, fig.height=22, message=FALSE, echo=FALSE, warning=FALSE}
Figure3
```

## Figure 4

Figure 4A

```{r message=FALSE, warning=FALSE}
# Variables of interest
variables_of_interest <- c("Risk_Score", 
                           "Prop_Day4_Alive", 
                           "Fs_Day4", 
                           "PP_Day4", 
                           "Fs_T", 
                           "PP_T")

# Create an empty dataframe to store p-values matrices
Data_stats_1 <- data.frame(
  variable = character(),
  OD_group = character(),
  Genotype = character(),
  Treatment = character(),
  `OD_group*Genotype` = character(),
  `OD_group*Treatment` = character(),
  `Genotype*Treatment` = character(),
  `OD_group*Genotype*Treatment` = character(),
  stringsAsFactors = FALSE
)
TableS3 <- list()

# Loop through each variable of interest
for (variable in variables_of_interest) {
  # Fit an ANOVA model
  model <- aov(as.formula(paste(variable, "~ OD_group * Genotype * Treatment")), data = Data)
  
  # Extract ANOVA table
  anova_table <- as.data.frame(anova(model))
  
  TableS3[[variable]] <- anova_table
  
  # Create a new row to append
  new_row <- data.frame(
    variable = variable,
    OD_group = anova_table$`Pr(>F)`[1],
    Genotype = anova_table$`Pr(>F)`[2],
    Treatment = anova_table$`Pr(>F)`[3],
    `OD_group*Genotype` = anova_table$`Pr(>F)`[4],
    `OD_group*Treatment` = anova_table$`Pr(>F)`[5],
    `Genotype*Treatment` = anova_table$`Pr(>F)`[6],
    `OD_group*Genotype*Treatment` = anova_table$`Pr(>F)`[7]
  )
  
  # Append the new row to Data_stats_1
  Data_stats_1 <- rbind(Data_stats_1, new_row)
}

Risk_Score_anova <- TableS3[["Risk_Score"]]
Prop_Day4_Alive_anova <- TableS3[["Prop_Day4_Alive"]]
Fs_Day4_anova <- TableS3[["Fs_Day4"]]
PP_Day4_anova <- TableS3[["PP_Day4"]]
Fs_T_anova <- TableS3[["Fs_T"]]
PP_T_anova <- TableS3[["PP_T"]]

Data_stats <- as.data.frame(melt(Data_stats_1, id.vars = c("variable"), variable.name = "vars"))

Data_stats$pval <- ifelse(Data_stats$value>0.05,"Not Significant",
                          ifelse(Data_stats$value<0.05 &
                                   Data_stats$value>0.01, "P<0.05",
                                 ifelse(Data_stats$value<0.01 &
                                          Data_stats$value>0.001, "P<0.01",
                                        ifelse(Data_stats$value<0.001, "P<0.001","NA"))))

Fig4A <- ggplot(Data_stats, aes(x=as.factor(vars),
                         y=as.factor(variable),
                         fill=as.factor(pval))) +
  geom_tile() +
  theme_bw() +
  labs(x="Factor",
       y="Measure of pathogenicity", 
       fill="p-value") +
  scale_fill_manual(values = c("white","maroon","red","pink")) +
  theme(axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=20, angle = 45, vjust = 1, hjust = 1),
        axis.title = element_text(size = 30),
        legend.position = "top",
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20))

```

**Figure 4A** P-values from ANOVA to indicate the significance of each factor corresponding to each measure of pathogenicity 

```{r fig.width=20, fig.height=20, message=FALSE, echo=FALSE, warning=FALSE}
Fig4A
```

Figure 4B

```{r message=FALSE, warning=FALSE}

# Select the columns for analysis
Corrdata <- Data[,c(27,34,42,46,48:50)]

# Compute the correlation matrix
M <- cor(Corrdata)

TableS4 <- as.data.frame(M)

# Compute the R^2 matrix by squaring the correlation coefficients
R2 <- M^2

# Correlation plot - Figure 4B
corrplot(R2, type = "upper", method = "color",
         order = "hclust",
         tl.col = "black", tl.srt = 45,
         cl.length = 11, cl.ratio = 0.2, col.lim = c(0,1))
```

**Figure 4B** Heatmap showing R2 (R=correlation coefficient) values to assess the differences between different measures of host susceptibility (survival, risk scores, and fraction symptomatic) and pathogen virulence (pathogen potential) 

[Prop_Day4_Alive: Proportion survival on Day 4 post-infection, PP_Day4: Pathogen potential on Day 4, PP_T: Pathogen potential on median survival day, Fs_T/T: Fraction symptomatic on median survival day/ Median survival day, Fs_Day4: Fraction symptomatic on Day 4, Fs_T: Fraction symptomatic on median survival day]. 


## Supplementary Figures

### Figure S1

```{r message=FALSE, warning=FALSE}
FigureS1 <- ggplot(LoadData, aes(x=OD,
                 y=log(`Bacterial_Load_CFU/fly`+1),
                 col=Treatment)) +
  geom_point(alpha=0.5,
             position = position_jitterdodge()) +
  stat_smooth(method = 'lm',
              se=FALSE) +
  theme_bw() +
  theme(axis.text=element_text(size = 24),
        legend.text = element_text(size = 20, face = "italic"),
        legend.position = "top",
        title=element_text(size = 20)) +
  labs(y="log bacterial load (CFU/fly)")
```

**Figure S1**- Bacterial load (CFU/fly) vs OD (A600nm) for different ODs and different pathogens. Single infected flies were homogenized in 500microL LB and plated on LB plates immediately after the infection.

```{r message=FALSE, echo=FALSE, warning=FALSE}
FigureS1
```

### Figure S2

```{r message=FALSE, warning=FALSE}

#Subsetting Data for PCA
Subset <- Data[,c(1:4,24:37,39:46,48:50)]

row.names(Subset) = paste(Subset$Genotype,Subset$Treatment,Subset$OD_group,Subset$Vial_ID)

PCA <- prcomp(t(Subset[,c(5:29)]), scale. = TRUE)
summary(PCA)

PCA_scores <- PCA$x

PCA_2 <- cbind(Subset,PCA_scores)

FigureS2 <- ggplot(as.data.frame(PCA_scores),aes(x=PC2,
                                     y = PC1,
                                     col = row.names(PCA_scores))) +
  geom_point(position = position_jitterdodge(),
             size = 3.5, alpha = 0.6) +
  theme_bw() +
  geom_label_repel(aes(label = row.names(PCA_scores)), nudge_x = -2, nudge_y = 2,
                   show.legend = FALSE) +
  theme(legend.position = "right",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 20)) +
  labs(x = "PC2: 23.66%",
       y = "PC1: 62.44%",
       col = "")
```

**Figure S2** Principal component analysis (PCA) over different measures of host susceptibility (survival, risk scores, and fraction symptomatic) and pathogen virulence (pathogen potential). [Prop_DayX_Alive: Proportion survival on Day X post-infection, PP_DayX: Pathogen potential on Day X post-infection, PP_T: Pathogen potential on median survival day, Fs_T/T: Fraction symptomatic on median survival day/ Median survival day, Fs_DayX: Fraction symptomatic on Day X post-infection, Fs_T: Fraction symptomatic on median survival day].

```{r fig.width=15, message=FALSE, echo=FALSE, warning=FALSE}
FigureS2
```

## Statistical Analysis

### Table S1

**Table S1:** Statistical analysis for analyzing the effect of OD and treatment (pathogen species) on log-transformed bacterial load. Data was fitted to a linear mixed-effects model. 

[Model: lm( log(`Bacterial_Load_CFU/fly`+1) ~ OD + Treatment)]

```{r message=FALSE, warning=FALSE}
Model1 <- lm(log(`Bacterial_Load_CFU/fly`+1)~OD+Treatment,data = LoadData)
summary(Model1)

Model2 <- lm(`Bacterial_Load_CFU/fly`~OD+Treatment,data = LoadData)
summary(Model2)

AIC(Model1, Model2)


MSum1 <- summary(Model1)
TableS1 <- as.data.frame(MSum1$coefficients)
print(TableS1)
```

### Table S2

**Table S2:** Statistical analysis for analyzing the effect of OD, treatment (pathogen species), and host genotype on survival. Data was fitted to a mixed-effects cox-proportional hazards model.

[Model: coxph(Survival ~ OD * Treatment * Genotype]

```{r message=FALSE, warning=FALSE}
cox <- coxph(Surv(Day_Dead, censor) ~ OD_group * Treatment * Genotype,
                  data = Data_1)
cox_summary <- summary(cox)
TableS2 <- as.data.frame(cox_summary$coefficients)


print(TableS2)

```


### Table S3

**Table S3:** ANOVA for analyzing the effect of OD, treatment (pathogen species), and host genotype on different measures of pathogenicity, corresponding to Figure 4A.

[Model: aov(“Measure of pathogenicity” ~ OD * Treatment * Genotype]

```{r message=FALSE, warning=FALSE}

# Variables of interest
variables_of_interest <- c("Risk_Score", 
                           "Prop_Day4_Alive", 
                           "Fs_Day4", 
                           "PP_Day4", 
                           "Fs_T", 
                           "PP_T")

# Create an empty dataframe to store p-values matrices
Data_stats_1 <- data.frame(
  variable = character(),
  OD_group = character(),
  Genotype = character(),
  Treatment = character(),
  `OD_group*Genotype` = character(),
  `OD_group*Treatment` = character(),
  `Genotype*Treatment` = character(),
  `OD_group*Genotype*Treatment` = character(),
  stringsAsFactors = FALSE
)
TableS3 <- list()

# Loop through each variable of interest
for (variable in variables_of_interest) {
  # Fit an ANOVA model
  model <- aov(as.formula(paste(variable, "~ OD_group * Genotype * Treatment")), data = Data)
  
  # Extract ANOVA table
  anova_table <- as.data.frame(anova(model))
  
  TableS3[[variable]] <- anova_table
  
  # Create a new row to append
  new_row <- data.frame(
    variable = variable,
    OD_group = anova_table$`Pr(>F)`[1],
    Genotype = anova_table$`Pr(>F)`[2],
    Treatment = anova_table$`Pr(>F)`[3],
    `OD_group*Genotype` = anova_table$`Pr(>F)`[4],
    `OD_group*Treatment` = anova_table$`Pr(>F)`[5],
    `Genotype*Treatment` = anova_table$`Pr(>F)`[6],
    `OD_group*Genotype*Treatment` = anova_table$`Pr(>F)`[7]
  )
  
  # Append the new row to Data_stats_1
  Data_stats_1 <- rbind(Data_stats_1, new_row)
}


Risk_Score_anova <- TableS3[["Risk_Score"]]
Prop_Day4_Alive_anova <- TableS3[["Prop_Day4_Alive"]]
Fs_Day4_anova <- TableS3[["Fs_Day4"]]
PP_Day4_anova <- TableS3[["PP_Day4"]]
Fs_T_anova <- TableS3[["Fs_T"]]
PP_T_anova <- TableS3[["PP_T"]]
```

#### A. Risk score

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(Risk_Score_anova)
```
#### B. Proportion alive on day 4

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(Prop_Day4_Alive_anova)
```
#### C. Pathogen potential on median survival day (PP_T)

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(PP_T_anova)
```
#### D. Pathogen potential on day 4 (PP_Day4)

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(PP_Day4_anova)
```
#### E.	Fraction symptomatic on median survival day (Fs_T)

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(Fs_T_anova)
```
#### F. Fraction symptomatic on day 4 (Fs_Day4)

```{r message=FALSE, warning=FALSE, echo=FALSE}
print(Fs_Day4_anova)
```


### Table S4

**Table S4:** Correlation matrix for different measures of pathogenicity containing R values corresponding to Figure 4B.

```{r message=FALSE, warning=FALSE}

# Select the columns for analysis
Corrdata <- Data[,c(27,34,42,46,48:50)]

# Compute the correlation matrix
M <- cor(Corrdata)

TableS4 <- as.data.frame(M)

print(TableS4)